### 2023-11-14 rmds/tables-supp.Rmd

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load results
source(here::here("code", "setup.R"))

data <- read_rds(here("data", "simulations-summarised.Rds"))
```

## Tables

Table S1.

```{r}
# Tidy up data, save, reorder manually, then reload
data %>% 
select(species, repro_freq, Nsamples, mesh_name_long, everything()) %>% 
  select(-sd_re, - sd_ae, -int_coverage, -mesh_name, -ntotal, -prop_pregnant, -n_maternal, -n_immature) %>% 
  filter(!par_name %in% "m95") %>% 
  pivot_wider(names_from = c("method_long", "par_name"), values_from = c("rel_error", "abs_error", "conv_success")) %>% 
  arrange(mesh_name_long, species, desc(repro_freq)) %>% 
  write_csv(here("tables", "table-s1.csv"))
```

Table S2. 
```{r}
data %>% 
  group_by(species, repro_freq, mesh_name, Nsamples, ntotal) %>% 
  nest() %>% 
  select(-data) %>% 
  ungroup() %>% 
  mutate(n = 250 - ntotal) %>% 
  filter(n > 0) %>% 
  write_csv(here("tables", "table-s2.csv"))

```