### 2023-11-14 rmds/tables.Rmd

---
output: 
  pdf_document:
    latex_engine: "pdflatex"
geometry: 
    - bottom=1.5cm
    - top=1.5cm
    - left=1.5cm
    - right=1.5cm
header-includes:
    - \fontsize{8}{20}
    - \usepackage{lscape}
    - \pagenumbering{gobble} 
tables: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source(here::here("code", "setup.R"))

# Load data
empirical <- read_rds(here("data", "empirical-summary.Rds"))
```

\
\
```{r tbl1, results = 'asis'}
t1 <- readxl::read_excel(here("data", "table-1.xlsx")) %>% 
  mutate(`School shark` = as.character(`School shark`)) %>% 
  mutate(`Gummy shark` = as.character(`Gummy shark`)) %>%  
  xtable(caption = "Descriptions and values of parameters used in a simulation study to evaluate the performance of alternative methods for estimating maternity parameters \\newline")

print(t1, type = "latex", caption.placement = "top", comment = F, include.rownames = F, 
      sanitize.text.function = identity, booktabs = TRUE)
```

\newpage

```{r tbl2, results = 'asis'}
t2 <- empirical %>% 
  rename(`$\\hat{P_{\\text{Max}}}$` = PMAX) %>% 
  rename(`$\\hat{L'_{50}}\\text{(cm)}$` = L50) %>% 
  rename(`$\\hat{L'_{95}}\\text{(cm)}$` = L95) %>% 
  rename(`$\\Delta_{i}$` = delta_AIC) %>% 
  rename(Method = method) %>%
  rename(`$w_i$` = w_AIC)

t2 <- xtable(t2, caption = "Parameter estimates and model selection criteria for three parameter logistic functions fit to empirical data for western North Atlantic sandbar sharks")

print(t2, type = "latex", caption.placement = "top", comment = F, include.rownames = F,
      sanitize.text.function = identity, booktabs = TRUE)
```